Illustrating Model Properties

### 2024-08-06 manuscript.qmd
add prevs of cIII depending on cII
fixes #9

>>>>

### Influence of Upstream Involvement

![Predicted (histograms) vs observed (beta posteriors) prevalence of scnearios that show how much more often the contralateral LNL III is involved when the upstream level II is involved (red) against when it is healthy (green). As in @fig-model-prevalences-with-ipsi, this is plotted for all four combinations of T-category and midline extension. Note that the contralateral LNL III is extremely rarely involved without the upstream level II.](figures/fig-model-prevalences-with-upstream.svg){#fig-model-prevalences-with-upstream}

Similarily to how we have shown the dependence of contralateral LNL II involvement on some extent of ipsilateral metastasis, we can look at the dependence of contralateral LNL III involvement depending on whether or not the upstream level harbors disease. In @fig-model-prevalences-with-upstream this is shown in analogy to @fig-model-prevalences-with-ipsi. Note that the prevalence of contralateral LNL III involvement is generally much lower than that of LNL II. Also, level III involvement is exceedingly rare when the upstream level II is not metastatic. The model captures this well.

### 2024-08-06 model.qmd
add prevs of cIII depending on cII
fixes #9

>>>>

## Contra LNL III Depending on Upstream LNL II Involvement

```{python}
import matplotlib.pyplot as plt
from matplotlib import gridspec
import numpy as np
from lyscripts.scenario import Scenario
from lyscripts.plot.utils import Histogram, BetaPosterior, draw, COLORS, split_legends
from scripts import paths, shared

full.set_title(
  "Observed vs Predicted Prevalence of Contralateral LNL III Involvement\nDependent on Upstream LNL III Involvement",
  pad=20.,
  fontsize="large",
)

for i, t_stage in enumerate(["early", "late"]):
  axes[i,0].set_ylabel(f"{t_stage} T-cat.", fontweight="bold")
  for j, midext in enumerate([False, True]):
    axes[0,j].set_title("Mid. ext." if midext else "Lateralized", fontweight="bold")
    axes[0,j].set_xticks([])
    axes[-1,j].set_xlabel("prevalence [%]")
    axes[i,j].set_yticks([])
    content = []
    for upstream, color in zip([False, True], [COLORS["green"], COLORS["red"]]):
      scenario = Scenario(
        t_stages=[t_stage],
        midext=midext,
        diagnosis={
          "contra": {"max_llh": {"II": upstream, "III": True}},
        }
      )
      content.append(
        Histogram.from_hdf5(
          filename=paths.model_dir / "full" / "prevalences_with_upstream.hdf5",
          dataname=scenario.md5_hash("prevalences"),
          color=color,
          label="prediction",
        )
      )
      content.append(
        BetaPosterior.from_hdf5(
          filename=paths.model_dir / "full" / "prevalences_with_upstream.hdf5",
          dataname=scenario.md5_hash("prevalences"),
          color=color,
        )
      )
    draw(
      axes=axes[i,j],
      contents=content,
      xlims=(0., 15. if midext else 4.),
    )

split_legends(
  axes=axes[0,0],
  titles=["upstream healthy", "upstream involved"],
  locs=[(0.4, 0.99), (0.5, 0.55)],
)
split_legends(
  axes=axes[0,1],
  titles=["upstream healthy", "upstream involved"],
  locs=[(0.3, 0.99), (0.6, 0.55)],
)
split_legends(
  axes=axes[1,0],
  titles=["upstream healthy", "upstream involved"],
  locs=[(0.2, 0.99), (0.65, 0.99)],
)
split_legends(
  axes=axes[1,1],
  titles=["upstream healthy", "upstream involved"],
  locs=[(0.3, 0.99), (0.65, 0.99)],
)

plt.savefig(paths.figure_dir / "fig-model-prevalences-with-upstream.svg")
plt.show()
```
